---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tauProcess

<!-- badges: start -->
<!-- badges: end -->

In clinical trials, the non-proportional hazard (NPH) scenarios are commonly encountered. The classical hazard ratio is no longer a meaningful treatment effect measure. Furthermore, the commonly used logrank test may lose its power under NPH. Several effect measures and testing procedures are proposed to overcome these problems, including weighted logrank, restricted mean survival time (RMST) and maxcombo tests. The local Kendall's tau measure we proposed is intuitive and clinically meaningful. The inference procedure based on local Kendall's tau is also provided. 

## Installation

You can install the development version of tauProcess from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("s07308/tauProcess")
```

## Usage & Example

This is a basic example which shows you how to fit a tau process:

```{r example}
library(tauProcess)
## basic example code
fit <- tau.fit(data = Inotuzumab)
```

You may use `summary()` to check the statistical inference results:
```{r}
summary(fit)
```

Furthermore, `plot()` will provide you the estimated tau process to investigate the evolution of treatment effect. 
```{r}
plot(fit)
```

