---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tauProcess

<!-- badges: start -->
<!-- badges: end -->

In clinical trials, the nonproportional hazard (NPH) scenarios are commonly encountered. Under NPH, the classical hazard ratio is no longer a meaningful treatment effect measure. Furthermore, the commonly used logrank test may lose its power. Several treatment effect measures and testing procedures are proposed to overcome these problems, including weighted logrank, restricted mean survival time (RMST) and maxcombo tests. The tau measure/process proposed is intuitive and clinically meaningful to accommodate the requirements from addendum to ICH E9 highlighting the interpretability of estimand. The inference procedure and plot based on tau measure and process are also provided. 

## Installation

You can install the development version of tauProcess from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("s07308/tauProcess")
```

## Usage & Example

This is a basic example which shows you how to estimate the tau process and make the corresponding statistical inference:

```{r example}
library(tauProcess)
fit <- tau.fit(data = pbc)
```

You may use `summary()` to check the inference results at the largest time specified:
```{r inference}
summary(fit)
```

Furthermore, `plot()` will provide you the estimated tau process to investigate the evolution of treatment effect: 
```{r plot-tau-process-no-cure}
plot(fit, type = "b")
```

For the case with possibly existing cure fraction, we may estimate the tau process for the susceptible subgroups:
```{r fit-cure}
fit_cure <- tau_proc(pbc, cure = TRUE)
```

The bootstrap method is recommended to do statistical inference about the tau process for susceptible subgroups. As the case with no cure fraction, we may plot the estimated tau process as well:
```{r plot-tau-process-with-cure}
plot(fit_cure, type = "b")
```


